---
title: Notes on matroid clustering experiments
---

```{r echo=F}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

## Execution time/quality tradeoff

This set of experiments eaims at exploring the relationship between
the running time of an algorithm and the radius of the returned clustering,
under different parameterizations.

The following algorithms are tested:
- `Random`: select a random maximal independent set as 
   the set of cluster centers.
- `ChenEtAl`: this is the baseline 7-approximation algorithm, which takes
  no parameters. We run it only on a sample of the full dataset.
- `SeqCoreset`: the sequential coreset construction, featuring as the only
  parameter $\tau$, which controls the size of the coreset
- `StreamCoreset`: constructs the coreset in streaming, with a single parameter
  $\tau$ controlling the coreset size, and hence the working memory
- `MRCoreset`: builds the coreset in parallel. Takes two parameters:
  $\tau$ which controls the size of the coreset, and the number of parallel
  workers to use, which ranges from 2 to 16.

For coreset-based algorithms, we set $\tau \in [512, 1024, 2048, 4096]$.
The number of outliers is fixed at $1\%$ of the size of the dataset.
We run on both the full Wikipedia dataset, and on a sample of 50\,000
points, which is the largest size that can be handled by
`ChenEtAl`^[which requires to pre-compute the distance matrix, thus requiring quadratic space].

As for the matroid constraint, the points in the dataset belong to one or more
of 100 categories, numbered from 0 to 99. 
In this first set of experiments, we impose a transversal matroid constraint
on the set of the first 10 categories.

The points in the dataset are derived from
Wikipedia pages using [GloVe](https://nlp.stanford.edu/projects/glove/).
As a distance metric we can use either the euclidean distance or the 
cosine distance.
In this set of experiments we consider both.

Each experiment is repeated 5 times, and the following plot
reports the averages.

```{r}
girafe(
    ggobj = drake::readd(plot_tradeoff),
    options = list(
        opts_sizing(width = 1)
    )
)
```
